### 2011-05-12 paper/paper.tex
Fixed proof for hit and run

Metropolis Markov chain methods, which include the basic ball-walking method, have two factors which cause increased runtime. The first and most theoretically important is that if the current point is close to the corner of the region, then the chance of a step succeeding can be exponentially small; a ball around that point might have most of its volume lying outside of the region $\mathcal R$. Another possibility is simply that, for instance, we may have a ball that it is so large it contains all of $\mathcal R$, but $\mathcal R$ is only a small fraction of the ball. In either case, there may be an exponentially small chance of a transition succeeding.

%Metropolis Markov chain methods, which include the basic ball walking method, have two major factors which can cause increased runtime. The first and more theoretically important is that if the current point is close to a corner of the region, then the chance of a step succeeding can be exponentially small. This arises even in simple cases such as a cube or a simplex if the sampling algorithm is given a point near a corner as the initial point. It also can occur if the step size is too large, the extreme case of which is sampling from a ball that is much larger than the entire admissible region.

%In this case shrinking the step size would clearly help, but it is not difficult to see that a small step size also mitigates the effect of nearby corners, as a larger fraction of the surrounding ball lies inside the admissible region.

The obvious solution is to simply decrease the radius of the ball, but this can cause another problem. If the step size is too small, it takes a long time to walk from one point in the admissible region to another. As it is certainly not possible for the Markov chain to mix before it is able to reach any other given point, a small step size causes the state space to have a large diameter, and thus a long mixing time.

We need to find a good balance between the two; that is, we cannot have the radius be too small or too large. Our attempt to find this balance is to introduce an adaptive walk, which we run for some time before running the standard ball walking or Metropolis method. This adaptive walk uses the success or failure of a step to guess whether its current step size is too small or too large, and then compensates by modifying it by a constant factor.

@@

Since the nonadaptive methods converge to the stationary distribution from any starting distribution, this yields a method that converges to the stationary distribution, but first attempts to both find a good step size in order to converge more rapidly and move to a better starting point (i.e. not near any corners).

@@

\begin{prop}
The uniform distribution is stationary under the hit-and-run method.
\end{prop}
\begin{proof}
Theorem \ref{symmetric} tells us that it suffices to show that $P(x,y) = P(y,x)$, where $P(x,y)$ is the probability distribution of transitioning from $x$ to $y$. Let $\ell$ be the line between $x$ and $y$. It is clear that in order to transition from $x$ to $y$, we must first choose $\ell$ as our line around $x$ and then choose $y$ as the point to transition to.

Therefore, we have $P(x,y) = P_x(\ell)P_\ell(y)$, and $P(y,x) = P_y(\ell)P_\ell(x)$, where $P_x$ is the probability distribution of lines starting from $x$ and $P_\ell$ is the probability distribution of points once we have chosen $\ell$. However, note that $P_x(\ell) = P_y(\ell)$, since we choose lines around a point uniformly, and that $P_\ell(y) = P_\ell(x)$, since once we have chosen $\ell$, we are choosing a point uniformly along it. Thus, $P(x,y) = P(y,x)$, and by Theorem \ref{symmetric}, the uniform distribution is stationary.
\end{proof}

>>>>

Metropolis Markov chain methods, which include the basic ball walking method, have two major factors which can cause increased runtime. The first and more theoretically important is that if the current point is close to a corner of the region, then the chance of a step succeeding can be exponentially small. This arises even in simple cases such as a cube or a simplex if the sampling algorithm is given a point near a corner as the initial point. It also can occur if the step size is too large, the extreme case of which is sampling from a ball that is much larger than the entire admissible region.

In this case shrinking the step size would clearly help, but it is not difficult to see that a small step size also mitigates the effect of nearby corners, as a larger fraction of the surrounding ball lies inside the admissible region.

The other contribution to the length of time required for Metropolis methods is that if the step size is too small, it takes a long time to walk from one point in the admissible region to another. As it is certainly not possible for the Markov chain to mix before it is able to reach any other given point, a small step size causes the state space to have a large diameter, and thus a long mixing time.

In order to find a good balance between the two main contributors to runtime stated above, we introduce an adaptive walk, which we run for some time before running the standard ball walking or Metropolis method. This adaptive walk uses the success or failure of a step to guess whether its current step size is too small or too large, and then compensates by modifying it by a constant factor.

@@

Since the nonadaptive methods converge to the stationary distribution from any starting distribution, this yields a method that converges to the stationary distribution, but also attempts to both find a good step size in order to converge more rapidly and move to a better starting point (i.e. not near any corners).

@@

\begin{prop}
The uniform distribution is stationary under the hit-and-run method.
\end{prop}
\begin{proof}
Theorem \ref{symmetric} tells us that it suffices to show that $P(x,y) = P(y,x)$, where $P(x,y)$ is the probability distribution of transitioning from $x$ to $y$. Let $\ell_{x,y}$ be the line between $x$ and $y$. It is clear that in order to transition from $x$ to $y$, we must first choose $\ell_{x,y}$ as our line around $x$ and then choose $y$ as the point to transition to.

As we are dealing with probability densities, we actually need to show that the probability of transitioning from $x$ to a neighborhood around $y$ is the same as the probability of transitioning from $y$ to a neighborhood around $x$ of the same size as that size goes to $0$.

To see this, we have $$P(x,N_y) = \int_{y' \in N_y}P_x(\ell_{x,y'})P_{\ell_{x,y'}}(y')\,dy$$ and $$P(y,N_x) = \int_{x'\in N_x}P_y(\ell_{x',y})P_{\ell_{x',y}}(x')\,dx'$$, where $P_x$ is the probability distribution of lines starting from $x$ and $P_\ell$ is the probability distribution of points once we have chosen $\ell$. However, note that $P_x(\ell_{x,y}) = P_y(\ell_{x,y})$, since we choose lines around a point uniformly, and that $P_{\ell_{x,y}}(y) = P_{\ell_{x,y}}(x)$, since once we have chosen $\ell$, we are choosing a point uniformly along it. As $N_x$ and $N_y$ get smaller, we have that $P_{\ell_{x,y'}}(y')$ and $P_{\ell_{x',y}}(x')$ both approach $P_{\ell_{x,y}}(x) = P_{\ell_{x,y}}(y)$ uniformly. Therefore, the difference between these two probabilities as the size of $N_x$ shrinks to zero approaches $P_{\ell_{x,y}}(x)\left(\int_{y' \in N_y}P_x(\ell_{x,y'})\,dy' - \int_{x' \in N_x}P_y(\ell_{x',y})\,dx'\right)$, and the two integrals are easily seen to be equal. Thus, $P$ is a symmetric distribution, and by Theorem \ref{symmetric}, the uniform distribution is stationary.
\end{proof}
